---
title: "purrring around with gapminder"
author: Sasha Lavrentovich 
output: html_notebook
---


```{r}
suppressMessages(library(tidyverse))
library(gapminder)
```


## Reacquaintance with gapminder 

We'll be looking at the relationship of life expectancy in 130-some countries on year (1950 through 2010). Here's a brute visual of that.  

```{r}
ggplot(gapminder, aes(x = year, y = lifeExp)) + 
  geom_jitter() + 
  geom_smooth(method = "lm")
```

We'll want to see if there is a linear relationship between life expectancy and year within each country so we'll want to fit one model to each country. Then we'll compare fits with comparing r-squared values and looking at residuals. 

## Fitting one model to many data frames 

### Data frames in data frames with tidyr::nest() 
First we'll create a nested dataframe where each country is a row and all the data about that country (lifeExp, pop, gdpPer) is put into a list with the nest() function. This will allow us keep models and data in correspondence with each other during oeprations. 

```{r}
# created new df grouped by country/continent and everything else nested 


# see one country, where you select the row and comma for all of the columns; see structure of new df 


# index into the dataframe's data column, and extract first element there, which turns out to be the data associated with Afghanistan 

```

### Create & apply a function with purrr::map()
We want to fit a model to each country in the dataframe. To do so, we'll create a function for a linear model fit for life expectancy and year, that we will then call for each country 

```{r}

# create a function for lm 


# fit one model 


# using dplyr::mutate and purrr:map, create a new column with the lm fit on each country's data row 
# fit one model on each country of the nested data frame


# each country has its own associated model now; use index to look at a specific country 


# use index to look into a specific country 


```

### Compute & collect model info with broom::tidy(), glance(), augment()
The following broom verbs can help with model information and diagnostics:  
1. tidy(): summarize the stats; returns tibble with model coefficients and stats  
2. glance(): see the stats; returns one-row of model summary  
3. augment(): add stats; returns tibble corresponding with original df, of observation-level stats 
Use broom to compute r squared values and collect residuals


```{r}
# model diagnostics as a one-line df

# model output as a tibble 

# model output related to original data points 

```

### Extract one statistic 
To measure linearity, we can compare r-squared values where if rsq is closer to 1 then the variance in life expectancy can be explained by year. We'll use glance() to see the r squared value for each country, and create a new column for that statistic. We'll also add the model outputs to our original data frame 

```{r}

# create new df with glance column, use that glance column to pull our the r.squared using map_dbl
# also add new tidy and augment columns for later work 


# see which countries have highest rsq; 


# you can see this plotted but its a lil overwhelming 
models_br %>% 
  ggplot(aes(rsq, reorder(country, rsq))) + 
  geom_point(aes(color = continent))

```

### Take out a df from your df with tidyr::unnest()
Say we want to collect the residuals and plot those out to see the quality of fit. Right now the residuals live within a dataframe within the dataframe, so to explore or perform operations on the model outputs, we can snap the lists back into data frames with unnest() 

```{r}

# unnest to original df; arguments: data, the column to unnest, .drop = true if you want to drop the other columns 

# unnest glance and drop the other columns 

# get residuals: unnest the augment summary, and plot the year and residuals by country, and facetted by continent 

```

### Other piping combinations  

```{r}

# can see which countries had non-linear progress by piping model outputs into other dplyr verbs 

models_br %>% 
  unnest(glance) %>% 
  filter(r.squared < 0.25)

# can also unnest by patterns such as string detections

models_br %>% 
  filter(str_detect(country, pattern = "Korea")) %>% 
  unnest(data)

# then plot it 

models_br %>% 
  filter(str_detect(country, pattern = "Korea")) %>% 
  unnest(data) %>% 
  ggplot() + 
    geom_line(aes(year, lifeExp, color = country))

```


## Now the opposite: fit many models to one data frame 

Here, we'll fit a couple different multiple linear regressions to all of the gapminder data (eh). We'll then extract the r-squared and AIC to compare the models. 
```{r}
# create a list of the different model combinations 
model_list <- list(
  simple = 'lifeExp ~ pop + gdpPercap', 
  medium = 'lifeExp ~ pop + gdpPercap + continent + year', 
  more = 'lifeExp ~ pop + gdpPercap + country + year', 
  woah = 'lifeExp ~ pop + gdpPercap + year*country')

# make model and their names into a data frame

model_frame <- data_frame(model = model_list) %>% 
  mutate(model_name = names(model))

# convert models into formulas 
model_frame <- model_frame %>% 
  mutate(model = map(model, as.formula))

# apply set of models to the data using map and pre-built lm function 



# check out the fit for one model 

```

### Diagnosing the regression

Since we have multiple regressions, we need to get the adjusted r-squared value since it accounts for the number of predictors we added (to avoid overfitting)


```{r}
# create a function that extracts the adj.r.squared value from a glance 


# add a rsq column that extracts the r.squared values for each model name 
# add an aic column that computes with the built-in AIC() function then arrange by AIC 


# next steps we could compare which predictor is more valuable than another with step wise comparison.. or cross-validate on fake data 

```

### Misc stuff 

```{r}

# find and convert factors: gapminder 'country' and 'continent' are factors but probably should be characters 

str(gapminder)

gapminder %>% 
  map_lgl(is.factor) # cycle through columns and identify if factor or not 

gapminder %>%  
  map_if(is.factor, as.character) %>% # if factor, then map as char 
  str()

# need to covnert back to data frame... 

```


## References 

Charlotte Wickham's tidyverse tutorial: https://github.com/rstudio/master-the-tidyverse 





