---
title: "Looking at tidytext"
output:
  html_document:
    df_print: paged
date: '2018-07-18'
---

## load libraries 

```{r}
library(gutenbergr)
library(tidyverse)
library(tidytext)
library(stringr)
```

## read in sherlock holmes stories from gutenberg library 

The gutenberg_download() uses the text ID number from the gutenberg library as its argument

```{r}
sherlock_raw <- gutenberg_download(1661)
```

## clean text

create new column that starts with all caps ADVENTURE, these are the short stories then add the text data to column, remove the main title, make story column into factors 

```{r}
sherlock <- sherlock_raw %>% 
  mutate(story = ifelse(str_detect(text, "ADVENTURE"), text, NA)) %>% 
  fill(story) %>% 
  filter(story != "THE ADVENTURES OF SHERLOCK HOLMES") %>% 
  mutate(story = factor(story, levels = unique(story)))

str(sherlock)
```

## tidy the text: add line number column, split text into word tokens, remove stop words, remove holmes      
```{r}
tidy_sherlock <- sherlock %>% 
  mutate(line = row_number()) %>% 
  unnest_tokens(word, text) %>% 
  anti_join(stop_words) %>% 
  filter(word != "holmes")

tidy_sherlock %>% 
  count(word, sort = TRUE)
```

## tf-idf: term frequency inverse document frequency 

```{r}
sherlock_tfidf <- tidy_sherlock %>% 
  count(story, word, sort = TRUE) %>% 
  bind_tf_idf(word, story, n) %>% 
  arrange(-tf_idf) %>% 
  group_by(story) %>% 
  top_n(10) %>% 
  ungroup() 

sherlock_tfidf

```
## visualize the tf-idf 

```{r}

sherlock_tfidf %>% 
  ggplot(aes(word, tf_idf, fill = story)) + 
  geom_col(alpha = 0.8, show.legend = FALSE) + 
  facet_wrap(~ story, scales = "free", ncol = 3) +
  coord_flip() + 
  theme(strip.text = element_text(size = 11)) + 
  labs(x = NULL, y = "tf-idf", title = "tf-idf in Sherlock Holmes")
```
## sentiment in sherlock holmes 

```{r}
sentiment <- tidy_sherlock %>% 
 # group_by(story)
  right_join(get_sentiments("nrc")) %>%
  filter(!is.na(sentiment)) %>% 
  count(sentiment, sort = TRUE)
sentiment
```
## plot the sentiment 

```{r}
sentiment_stories <- tidy_sherlock %>% 
  group_by(story) %>% 
  mutate(word_count = 1:n(), 
         index = word_count %/% 500 + 1) %>% 
  inner_join(get_sentiments("bing")) %>% 
  count(story, index = index, sentiment) %>% 
  ungroup() %>% 
  spread(sentiment, n, fill = 0) %>% 
  mutate(sentiment = positive - negative)

sentiment_stories %>% 
  ggplot(aes(index, sentiment, fill = story)) + 
  geom_bar(alpha = 0.5, stat = "identity", show.legend = FALSE) + 
  facet_wrap(~ story, ncol = 22, scales = "free_x")

sentiment_stories
```


